Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-14 manuscript/bef_stability/manuscript.Rmd
Last edits before submission

- Number of words:
  - Abstract: 172
  - Main text: 5383

- Number of references: 87 
- Number of figures: 2
- Number of tables: 1
- Number of text boxes: 0

@@

Both biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships have received considerable attention during the last decades.
Still, the BEF relationship remains rarely investigated in the context of food
webs, especially in empirical studies explicitly considering foodweb structures.
Foodweb stability-complexity relationship has been mainly the focus of
theoretical studies with very little empirical testing. To fill this gap, we linked
species richness, foodweb structure, annual biomass and its temporal
stability using a standardized monitoring of fish communities in French streams
running from 1995 to 2018. We inferred trophic interactions from
individual body size and diet literature. We reveal that species richness and average
trophic level had a positive relationship with annual biomass of fish
communities while they had respectively a negative and positive effect on its
temporal stability. We also found that the effects of stream size and
temperature on annual biomass and stability are mediated by their effects on
diversity and foodweb structure. Overall, our study opens fruitful perspectives
on the link among species richness, network structure, community biomass and its
temporal stability.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006].
Species richness has been shown to promote higher biomass production and  
higher temporal stability of total biomass in a variety of ecosystems
and taxa. 
Indeed species richness generally increases primary productivity and
total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders, [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. 
In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020].
Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning in experimental and empirical studies remains mostly focused
on single trophic levels, ignoring the effects of foodweb structure. 
This gap is even more striking for the relationship between diversity and
temporal stability of ecosystem functioning, in spite of long-standing interest
for the links between food web complexity and stability in theoretical ecology
[e.g. @may_will_1972; @dunne_network_2006; @neutel_reconciling_2007;
@duffy_why_2009; @stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and foodweb structure should modulate ecosystem propertiestotal
biomass and their stability of communities. Complementarity in the use of
resources, a main mechanism previously highlighted in single-trophic studies, is
important to understand the effects of species richness in food webs [@thebault_food-web_2003;
@poisot_trophic_2013; @schneider_animal_2016], but in that case it also depends
on food web connectance [@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in food webs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can in some cases reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, recent models revealed new mechanisms by
which food web structure affects total ecosystem biomass and its temporal
variability. In particular, the number of trophic levels in food webs, which is
associated with the presence of species at higher trophic levels, increases
total primary production [@wang_biodiversity_2018]. The addition of predators has
also cascading effects on both species variability and synchrony at lower
trophic levels [@teng_dynamics_2004; @shanafelt_stability_2018].
Overall, these theoretical studies show that the effects of species richness and
network structure, such as connectance and average trophic level, are
potentially conflicting or synergistic in such a way that they need to be
addressed jointly.

The gap in studies that consider jointly effects of diversity and food web
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and food web complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that describing
trophic interactions between species is highly labour intensive. These
challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires time-series data and longer-term
analyses. In this context, recent developments to predict species trophic
interactions [@gravel_inferring_2013], combined with observational long-term surveys
of ecological communities, can offer an opportunity to investigate the links
between diversity, food web structure and the stability of total community
biomass in natural settings.

Such approach based on data from natural communities can in addition give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@de_boeck_patterns_2017; @hansen_climate_2013;
@bluthgen_land_2016; @oliver_heterogeneous_2010; @franssen_annual_2011;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017]. Environment also
directly affect species richness and food web structure. For instance, larger
ecosystems host food webs with higher diversity and trophic levels
[@mchugh_dual_2010] while temperature can affect food web diversity and
structure in contrasted ways depending on the ecosystem studied
[@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing data
from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. Despite that fish trophic
interactions are difficult to measure in natural conditions, a substantial body
of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Especially in stream communities, including fishes,
@brose_predator_2019 found that the prediction of predator prey trophic
interactions based on body size ratio was far better in stream ecosystems than
in other aquatic and terrestrial ecosystems. Lastly, most of the
diversity-biomass and diversity-stability relationships, have been assessed in
terrestrial ecosystems while there is a evident need to test such relationships
in a variety of ecosystem types.

Our aim is to document the relationships among species richness, foodweb structure, 
total biomass and its temporal stability in natural communities. To do so, we
used the ONEMA dataset consisting in standardized sampling of fish communities
in French streams [@edeline_ecological_2013]. We selected 99 sites that were sampled for more than 10
years spanning on the 1995-2018 period. We inferred 1496 interaction networks,
based on body size and diet. We question whether species richness, connectance
and average trophic level increase, decrease or are not related to
total biomass and its temporal stability. We further question whether the
species richness, connectance, and average trophic level have antagonistic or
mutual effect on total biomass and its temporal stability. Finally, we
question if ecosystem size and average temperature increase total biomass and
its temporal stability.

@@

```{r}
myload(qt_station, dir = mypath("data"))
matmet_cap <- paste0( 
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. The red dots show the location of (C) and (D) sites.
  ", " Inner borders draw the limits of the hydrographic basins. ", "(B) The
  metaweb of the study. Each trophic species (N = 412) is represented by a node
  and each
  color represent a species (S = 45). Resource nodes (7) were drawn in grey. ",
  "The species names corresponding to the three digit codes are found in Table
  S1." ,
  "Yearly temporal series of networks of two stations having respectively a low
  (C) and high (D) temporal stability of biomass. C and D stations have
  respectively biomass stability the closest to the first and third quartile of
  the distribution of the temporal stability of biomass in the dataset,
  respectively ", round(qt_station$bm_std_stab[1],2), " and ",
  round(qt_station$bm_std_stab[2],
    2), ". Point size are proportional to the biomass of the trophic species.")

@@

```{r matmet, fig.cap = matmet_cap, results = "fig.show", out.height="80%"}
library(knitr)
include_graphics(path = mypath("manuscript", "bef_stability", "figs", "fig1.pdf"))
```

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in Supplementary material, part 1.1). For streams
shallower than 0.7m across the entire width, the sampling took place over the
whole width of the streams and was done by foot whereas in streams deeper than
0.7m, the sampling was done over the bank by foot or by boat. The sampling
season spanned from late spring to autumn. One sampling event correspond to a
sampling a given year in a given sampling station. During sampling, all the
fishes caught were identified. For each species, either the body size of all
fish was measured, or a representative sample when the number of fish was high
(Figure S1, supplementary information). The fishes were released at the end of
the sampling. Additionally, the physical characteristics of the stream were 
recorded during sampling event as well as the sampling effort, quantified as the
surface sampled (range `r range_surface[1]`-`r range_surface[2]` square metres).

The stations where the stream was of intermediate size could be sampled
alternatively with one of the two sampling protocols. In this case, we
kept only the most frequently used sampling protocol. Then, we selected the
stations that had been sampled ten years or more. There was low heterogeneity of
the sampling month within stations, with the median sampled month being
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$. As the concept of community variability referred to
variability around an equilibrium, which is rarely observed or difficult to
characterize in natural communities, only stations with no temporal trends
in total biomass were selected (detailed in supplementary methods). However, the
results were robust to the relaxation of this hypothesis (Figure S7). These
selection steps resulted in 99 stations distributed over seven hydrographic
basins (Figure \@ref(fig:matmet) A).

## Variable computation

@@

The body size of all fish individuals was not recorded when they were too
numerous. The body size of the unmeasured individuals was inferred under the
assumption that species body size followed a normal distribution whose mean and
standard deviation were quantified from the recorded individuals of the sampling
(See supplementary methods, Figure S1). The body mass of each individual was
then estimated using an allometric law $B_i = 0.01 \times l_i^{3.03}$, $B$
with the body mass in grams, $l$ the body size in millimeters, $i$ the
individual and where the coefficients $0.01$ and $3.03$ comes from two
meta-analyses on fishes [@lleonart_removing_2000; @froese_cube_2006]. Species
biomass and total biomass at each sampling event was then calculated by summing
individual body masses. To account for the variation in sampling effort among
sites, species biomass and total biomass have been related to the surface
sampled, i.e. $gm^{-2}$. In the analysis, the total biomass of a
station has been defined as the median of the total biomass across sampling
events.

@@

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

$$
\phi_i = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sqrt{\sigma^2_{x_{k,i}}}})^2}
$$

@@

Species richness was computed as the total number of species present across
sampling events 
in a given station. As for biomass, we controlled species richness for sampling
effort by dividing the total number of species by the total surface sampled in
the given station. Species richness was then expressed in number of species per 
square meter sampled.

@@

The trophic interaction network was inferred from body size and ontogenic diet shift
following previous studies [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as for many organisms, shifts
throughout an individual's life, and thereby with body size. For example, alvins
feed on plankton while adults may feed on algae and fishes, with bigger individuals
eating bigger
fishes. Each fish species was hence divided in nine body size classes, each
corresponding to a trophic species (Figure S2). Additionally to the fish trophic
species, we added seven resource nodes present in the ontogenic food diet database:
detritus, biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and
zoobenthos (grey nodes in Figure \@ref(fig:matmet) B, C & D). The fish trophic
species and the resources constituted the 412 nodes of the metaweb, 45 fish
species divided into nine body size classes plus seven resource nodes. The inference
of trophic interactions was shown to be more accurate when considering body size
classes, i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. Species identity allowed to
differentiate between piscivorous and non piscivorous species, to know which
resources the trophic species are feeding on and to know the minimum size at
which a piscivorous species begin to feed on fishes. The ontogenic fish diet
database was filled according to fishbase and literature (Table S2). Each
fish species had two or three life stages which are delimited by a body size
range (Figure S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Figure S2).

For a piscivorous trophic species, the fish-fish
trophic links was defined in two steps and based on the body size ratio between
predators and prey. First of all, the predation window of a piscivorous trophic specie was
defined as 3% to 45% of its midpoint [@mittelbach_ontogeny_1998;
@claessen_impact_2002; @hart_handbook_2008]. Then, a trophic link was set
between the piscivorous trophic species and every trophic species whose midpoint
was included in the predation window (Figure S2). The fish-resource trophic
links were set between the trophic species and the resources nodes according the
food diet of their size class, for both the piscivorous and non piscivorous
trophic species. Lastly, the trophic links among resource nodes were set
according to the literature [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and year, we
computed two of the most commonly used network metrics to characterize food
webs, namely the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the maximum number of interactions if all species interact
together ($C = L / N^2$, L and N being resp. the number of links and of trophic
species plus resource nodes). The mean trophic level $\overline{T}$ was computed
as the mean trophic level weighted by the biomass of each trophic species
(excluding resource nodes for which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, with $T_i$ and
$B_i$ being respectively the trophic level and the biomass of the node $i$. The
trophic level of the nodes was computed as one added to the average trophic
level of its food item. For each site, the median of connectance and average
trophic level over time was used in subsequent analysis.

@@

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow. As data from the naiades database are
not collected on the same site as our fish sampling sites, we performed an
interpolation with spatial stream networks [@isaak_applications_2014;
@hoef_ssn:_2014]. The interpolation procedure is detailed in supplementary (part
1.2).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of for water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigen values distribution (Figure S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Figure S4). A varimax axis rotation was
performed to maximize the representation of the variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. The
second axis represented a gradient of increasing average water temperature and
decreasing altitude and slope. Further description of the variables are provided
in supplementary information (part 1.3).

@@

Structural Equation Modeling enables us to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. The structural equation
model explaining the annual total biomass allowed the environment
to affect directly species richness and network structure, as well as annual total
biomass (Figure \@ref(fig:sem-plot) A). We assumed that both species richness
and network structure directly affect total biomass (Figure \@ref(fig:sem-plot)
A) and that species richness affected foodweb structure
[@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010].

@@

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and the manuscript are accessible on github. The analysis were
performed with R version 3.6.3 [@r_core_team_r_2020].

@@

#ctrl <- lmeControl(opt='optim')
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
```

>>>>

- Running title: Foodweb structure, biomass and stability

- Number of words:
  - Abstract: 151
  - Main text: 5535

- Number of references: 92
- Number of figures: 2
- Number of tables: 1
- Number of text boxes: 0

@@

Biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships are central to ecology. However, they remain largely untested in a
foodweb context in natural settings. Here we study the links among
environmental conditions, richness, foodweb structure, annual biomass and its
temporal stability using a standardized monitoring of 99 stream fish communities
spanning from 1995 to 2018. We first reveal that both richness and average
trophic level are positively related to annual biomass, with effects of similar
strength. Second, we show that network structure affects community stability,
with a positive effect of mean trophic level, and contrary to general
expectation, we find that richness destabilizes fish communities. Finally, we
find that environmental conditions affect both biomass and its stability via
effects on richness and network structure. Strikingly, effects on stability are
mediated by changes in species stability rather than synchrony, which contrasts
with results from single-trophic communities and might be characteristic of
multi-trophic communities.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006].  Species richness has
been shown to promote higher biomass production and  higher temporal stability
of total biomass in a variety of ecosystems and taxa.  Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], but
also in briophyte communities [@mulder_physical_2001], stream feeders,
[@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018].  In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020].  Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning in experimental and empirical studies remains mostly focused on
single trophic levels, ignoring the effects of foodweb structure.  This gap is
even more striking for the relationship between diversity and temporal stability
of ecosystem functioning, in spite of long-standing interest for the links
between foodweb complexity and stability in theoretical ecology [e.g.
@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and foodweb structure should modulate ecosystem properties such as
total biomass and its stability. Complementarity in the use of
resources, a main mechanism previously highlighted in single-trophic studies, is
important to understand the effects of species richness in foodwebs [@thebault_food-web_2003;
@poisot_trophic_2013; @schneider_animal_2016], but in that case it also depends
on foodweb connectance [@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in foodwebs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can in some cases reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, recent models revealed new mechanisms by
which foodweb structure affects total ecosystem biomass and its temporal
variability. In particular, the number of trophic levels in foodwebs, which is
associated with the presence of species at higher trophic levels, increases
total primary production [@wang_biodiversity_2018]. The addition of predators has
also cascading effects on both species variability and synchrony at lower
trophic levels [@teng_dynamics_2004; @shanafelt_stability_2018].
Overall, these theoretical studies show that the effects of species richness and
network structure, such as connectance and average trophic level, are
potentially conflicting or synergistic in such a way that they need to be
addressed jointly.

The gap in studies that consider jointly effects of diversity and foodweb
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and foodweb complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that describing
trophic interactions between species is highly labour intensive. These
challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires time-series data and longer-term
analyses. In this context, recent developments to predict species trophic
interactions [@gravel_inferring_2013], combined with observational long-term surveys
of ecological communities, can offer an opportunity to investigate the links
between diversity, foodweb structure and the stability of total community
biomass in natural settings.

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing
data from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allows for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. Despite that fish trophic interactions are
difficult to measure in natural conditions, a substantial body of literature
supports that they can be inferred through body size and diet informations
[@jennings_weak_2001; @gravel_inferring_2013; @brose_predator_2019]. Especially
in stream communities, including fishes, @brose_predator_2019 found that the
prediction of predator-prey trophic interactions based on body size ratio was
far better in stream ecosystems than in other aquatic and terrestrial
ecosystems. Lastly, most of the diversity-biomass and diversity-stability
relationships, have been assessed in terrestrial ecosystems while there is a
evident need to test such relationships in a variety of ecosystem types.

Our aim is to document the relationships among species richness, foodweb
structure, total biomass and its temporal stability in natural communities. To
do so, we used the ONEMA dataset consisting in a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years spanning on the 1995-2018 period. We
inferred 1496 interaction networks, based on body size and diet. We question
whether species richness, connectance and average trophic level increase,
decrease or are unrelated to total biomass and its temporal stability. We
further question whether the species richness, connectance, and average trophic
level have antagonistic or mutual effect on total biomass and its temporal
stability. Finally, we question if ecosystem size and average temperature
increase total biomass and its temporal stability.

@@

```{r, cache = FALSE}
myload(qt_station, dir = mypath("data"))
matmet_cap <- paste0( 
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. The red dots show the location of (C) and (D) sites.
  ", " Inner borders draw the limits of the hydrographic basins. ", "(B) The
  metaweb of the study. Each trophic species (N = 412) is represented by a node
  and each
  color represents a species (S = 45). Resource nodes (7) are drawn in grey. ",
  "The species names corresponding to the three digit codes are found in Table
  S1." ,
  "Yearly temporal series of networks of two stations having respectively a low
  (C) and high (D) temporal stability of biomass. C and D stations have
  respectively biomass stability the closest to the first and third quartile of
  the distribution of the temporal stability of biomass in the dataset,
  respectively ", round(qt_station$bm_std_stab[1],2), " and ",
  round(qt_station$bm_std_stab[2],
    2), ". Point size is proportional to the biomass of the trophic species.")

@@

```{r matmet, cache = FALSE, fig.cap = matmet_cap, results = "fig.show", out.height="80%"}
library(knitr)
include_graphics(path = mypath("manuscript", "bef_stability", "figs", "fig1.pdf"))
```

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). For streams
shallower than 0.7m across the entire width, the sampling took place over the
whole width of the streams and was done by foot whereas in streams deeper than
0.7m, the sampling was done over the bank by foot or by boat. The sampling
season spanned from late spring to autumn. One sampling event corresponds to a
sampling a given year in a given sampling station. During sampling, all the
fishes caught were identified. The fishes were released at the end of
the sampling. Additionally, the width and the depth of the stream were 
recorded during sampling event as well as the sampling effort, quantified as the
surface sampled (range `r range_surface[1]`-`r range_surface[2]` square metres).

The stations where the stream was of intermediate size could be sampled
alternatively with one of the two sampling protocols. In this case, we
kept only the most frequently used sampling protocol. Then, we selected the
stations that had been sampled ten years or more. There was low heterogeneity of
the sampling month within stations, with the median sampled month being
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$. As the concept of community variability referred to
variability around an equilibrium, which is rarely observed or difficult to
characterize in natural communities, only stations with no temporal trends
in total biomass were selected (detailed in appendix B, part 2). However, the
results were robust to the relaxation of this hypothesis (Figure S7). These
selection steps resulted in 99 stations distributed over seven hydrographic
basins (Figure \@ref(fig:matmet) A).

## Total biomass and its temporal stability

@@

The body size of all fish individuals was not recorded when they were too
numerous (Figure S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution whose mean and standard deviation were quantified from the
recorded individuals of the sampling (Figure S1, appendix B, part 1.2). The body
mass of each individual was then estimated following the results of two
meta-analyses on fishes [@lleonart_removing_2000; @froese_cube_2006], using an
allometric law $B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams,
$l$ the body size in millimeters, $i$ the individual and where the coefficients
$0.01$ and $3.03$.  Species biomass and total biomass at each sampling event
were then calculated by summing individual body masses. To account for the
variation in sampling effort among sites, species biomass and total biomass were
corrected by the surface sampled, and expressed in grams by square meter ($gm^{-2}$). In the analysis,
the temporal average of the annual total biomass of a station has been defined
as the median of the total biomass across sampling years.

@@

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

$$
\phi_i = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sqrt{\sigma^2_{x_{k,i}}}})^2} = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sigma_{x_{k,i}}})^2}
$$

## Species richness and foodweb structure

@@

Species richness was computed as the total number of species present across
all the sampling events in a given station. As for biomass, we controlled
species richness for sampling effort by dividing the total number of species by
the total surface sampled in the given station. Species richness was then
expressed in number of species per square meter sampled.

@@

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Figure S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Figure \@ref(fig:matmet) B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. Species identity allowed to
differentiate between piscivorous and non piscivorous species, to know which
resources the trophic species are feeding on and to know the minimum size at
which a piscivorous species begin to feed on fishes. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by a body size
range (Figure S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Figure S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic specie was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Figure S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class, for both the piscivorous and non piscivorous trophic species.
Lastly, the trophic links among resource nodes were set according to the
literature [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and year, we
computed two of the most commonly used network metrics to characterize food
webs, namely the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the maximum number of interactions if all species interact
together ($C = L / N^2$, L and N being resp. the number of links and of trophic
species plus resource nodes). The mean trophic level $\overline{T}$ was computed
as the mean trophic level weighted by the biomass of each trophic species
(excluding resource nodes for which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and
$B_i$ being respectively the trophic level and the biomass of the node $i$. The
trophic level of the nodes was computed as one added to the average trophic
level of its food item. For each site, the median of connectance and average
trophic level over time was used in subsequent analysis.

@@

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow. As data from the naiades database are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Figure S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Figure S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, DBO and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure as measured by connectance and
average trophic level [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Figure
\@ref(fig:sem-plot) A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

@@

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and the manuscript are accessible on github (alaindanet/fishcom). The analyses were
performed with R version 3.6.3 [@r_core_team_r_2020].

@@

#ctrl <- lmeControl(opt='optim' relative)
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
```
